---
title: "Reproducible Research-Programming Assignment I"
author: "Marlen Mueller"
date: "5/8/2017"
output: md_document
---

# Reading in the dataset
```{r Readin in data}
setwd("~/Desktop/CourseraDataScience/")
activity_data <- read.csv("activity 2.csv", header=TRUE, stringsAsFactors = FALSE)
```
Variables:   

 * steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
 
 * date: The date on which the measurement was taken in YYYY-MM-DD format
 
 * interval: Identifier for the 5-minute interval in which measurement was taken

# What is mean total number of steps taken per day?
Histogram of the total number of steps taken each day
```{r}
activity_data$date <- as.Date(activity_data$date)
```
calculate the total number of steps per day
```{r}
steps_per_day <- tapply(activity_data$steps, activity_data$date, sum)
```
plot the histogram
```{r histogram_with_NAs}
hist(steps_per_day, breaks = length(steps_per_day), xlab = "Steps per Day", main = "Histogram of the total number of steps taken each day")
```

# Mean and median number of steps taken each day
calculate the mean of steps taken per day (including missing values)
```{r}
mean(steps_per_day, na.rm = TRUE)
```
calculate the median of steps taken per day (including missing values)
```{r}
median(steps_per_day, na.rm = TRUE)
```
take out missing values in steps and the entire data frame
```{r}
activity_full <- activity_data[!is.na(activity_data$steps), ]
```

# What is the average daily activity pattern?
averaged steps per interval
```{r}
averageInterval <- tapply(activity_full$steps, activity_full$interval, mean)
```
```{r daily_activity_with_NAs}
plot(as.numeric(names(averageInterval)), averageInterval, type="l", xlab = "Intervals", ylab = "Average Steps")
```
The 5-minute interval that, on average, contains the maximum number of steps
```{r}
names(averageInterval[which.max(averageInterval)])
```
# Imputing missing values
replace missing values in variable steps with mean of all steps
```{r}
activity_data[is.na(activity_data)] <- mean(activity_data$steps, na.rm = TRUE)
```
Histogram of the total number of steps taken each day after missing values are imputed
```{r histogram_without_NAs}
steps_per_day <- tapply(activity_data$steps, activity_data$date, sum)
hist(steps_per_day, breaks = length(steps_per_day), xlab = "Steps per Day", main = "Histogram of the total number of steps taken each day", sub = "with missing values imputed")
```
# Are there differences in activity patterns between weekdays and weekends?
add a column to the data frame "activity_data" displaying the weekdays
```{r}
activity_data$days <- weekdays(activity_data$date)
```
filter 2 data frames for weekdays and weekends
```{r}
weekdays_data <- activity_data[(activity_data$days != "Saturday") & (activity_data$days != "Sunday"), ]
weekends_data <- activity_data[(activity_data$days == "Saturday") & (activity_data$days == "Saturday"), ]
```
take the mean of the steps per interval for each of the newly generated data frames
```{r}
weekdays_ave <- tapply(weekdays_data$steps, weekdays_data$interval, mean)
weekends_ave <- tapply(weekends_data$steps, weekends_data$interval, mean)
```
make a panel plot comparing the average number of steps taken per 5 minute interval across weekdays and weekends
```{r planel_plot}
par(mfrow=c(2,1))
plot(as.numeric(names(weekdays_ave)), weekdays_ave, type="l", xlab = "Intervals", ylab = "Average Steps", main = "Average Steps on Weekdays")
plot(as.numeric(names(weekends_ave)), weekends_ave, type="l", xlab = "Intervals", ylab = "Average Steps", main = "Average Steps on Weekends")
```
